%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\baselineskip = 1.0\normalbaselineskip
\begin{thebibliography}{HMRR93}

\bibitem{andreou2002radiocoloring}
M.~I. Andreou, D.~A. Fotakis, S.~E. Nikoletseas, V.~G. Papadopoulou, and P.~G.
  Spirakis.
\newblock On radiocoloring hierarchically specified planar graphs:
  Pspace-completeness and approximations.
\newblock In {\em Mathematical Foundations of Computer Science 2002: 27th
  International Symposium, MFCS 2002, Warsaw, Poland, August 26-30, 2002.
  Proceedings}, volume 2420, page~81. Springer Science \& Business Media, 2002.


\bibitem{bottcher2020simulation}
S.~B{\"o}ttcher, R.~Hartel, and S.~Peeters.
\newblock Simulation computation in grammar-compressed graphs.
\newblock {\em arXiv preprint arXiv:2001.04760}, 2020.


\bibitem{brenguier2012comparison}
R.~Brenguier, S.~G{\"o}ller, and O.~Sankur.
\newblock A comparison of succinctly represented finite-state systems.
\newblock In {\em International Conference on Concurrency Theory}, pages
  147--161. Springer, 2012.

\bibitem{chen2003periodic}
H.~Chen.
\newblock Periodic constraint satisfaction problems: Polynomial-time
  algorithms.
\newblock In {\em International Conference on Principles and Practice of
  Constraint Programming}, pages 199--213. Springer, 2003.

\bibitem{chen2005periodic}
H.~Chen.
\newblock Periodic constraint satisfaction problems: Tractable subclasses.
\newblock {\em Constraints}, 10(2):97--113, 2005.

\bibitem{goller2005fixpoint}
S.~G{\"o}ller and M.~Lohrey.
\newblock Fixpoint logics on hierarchical structures.
\newblock In {\em International Conference on Foundations of Software
  Technology and Theoretical Computer Science}, pages 483--494. Springer, 2005.

 \bibitem{lohrey2012model}
M.~Lohrey.
\newblock Model-checking hierarchical structures.
\newblock {\em Journal of Computer and System Sciences}, 78(2):461--490, 2012.

\bibitem{marathe1995complexity}
M.~V. Marathe, H.~B. Hunt~III, R.~E. Stearns, and V.~Radhakrishnan.
\newblock Complexity of hierarchically and 1-dimensional periodically specified
  problems.
\newblock Technical report, Los Alamos National Lab., NM (United States), 1995.


%%  

\bibitem[AC94]{AC94} S. Agarwal and  A. Condon, 
``On Approximation Algorithms for Hierarchical MAX-SAT,''
{\em Proc of the 10th IEEE Conference on Structures in Complexity Theory,} 
July, 1995.


\bibitem[AC+95]{AC+95} S. Agarwal, A. Condon, 
M.V. Marathe, H.B. Hunt III and R.E. Stearns 
``On Approximation Algorithms for Periodically and Hierarchically Specified
Problems,'' manuscript under preparation,
August, 1995.


\bibitem[BOW83]{BOW83} J.L. Bentley, T. Ottmann, P. Widmayer,
``The Complexity of Manipulating Hierarchically Defined set of Intervals,''
{\em Advances in Computing Research, ed.  F.P. Preparata} 
Vol. 1, (1983), pp. 127-158.





\bibitem[CM93]{CM93} 
E. Cohen and  N. Megiddo,
``Strongly Polynomial-time and NC Algorithms for 
Detecting Cycles in Dynamic Graphs,''
{\em Proc. 21st ACM Annual Symposium on Theory of Computing (STOC)}, 
1989, pp. 523-534. 
Journal version  appears in {\em Journal of the ACM (J. ACM)} Vol. 40,
No. 4, September 1993, pp. 791-830.




\bibitem[CM91]{CM91} 
E. Cohen and N. Megiddo,
``Recognizing Properties of Periodic graphs'',
{\em Applied Geometry and Discrete Mathematics, The Victor Klee Festschrift}
Vol. 4, P. Gritzmann and B. Strumfels, eds., ACM, New York, 1991, pp. 135-146.




\bibitem[CF+93]{CF+93} A. Condon, J. Feigenbaum, C. Lund and P. Shor,
``Probabilistically Checkable Debate Systems and Approximation Algorithms
for PSPACE-Hard Functions'', in  
{\em Proc.  25th ACM Symposium on Theory of Computing (STOC)}, 
1993, pp. 305-313.


\bibitem[CF+94]{CF+94} A. Condon, J. Feigenbaum, C. Lund and P. Shor,
``Random Debators and the Hardness of Approximating
Stochastic functions for PSPACE-Hard Functions,'' 
{\em Proc. 9th IEEE Annual Conference on Structure in Complexity Theory}, 
June 1994, pp. 280-293. 


\bibitem[DP60]{DP}
M.Davis and H.Putnam, 
``A Computing Procedure for Quantification Theory,''
{\em Journal of the ACM (J. ACM)} 
Vol. 7, 1960, pp. 151-158.




\bibitem[DLR79]{DLR79}
D. Dobkin, R.J. Lipton and S. Reiss,
``Linear programming is log-space hard for P'',
{\em Information Processing Letters (IPL)} 
Vol. 8, 1979, pp. 96-97.


\bibitem[FF58]{FF58} 
L.R. Ford and D.R. Fulkerson,
``Constructing Maximal Dynamic Flows from Static Flows,''
{\em Operations Research,}
No. 6, 1958, pp. 419-433.




\bibitem[Ga82]{Ga82} H. Galperin 
``Succinct Representation of Graphs,''
Ph.D. Thesis, Princeton University, 1982.





\bibitem[GW83]{GW83} 
H. Galperin and A. Wigderson,
``Succinct Representation of Graphs,''
{\em Information and Control }, 
Vol. 56, 1983, pp. 183-198.






\bibitem[Ga59]{Ga59} 
D. Gale,
``Transient Flows in Networks,''
{\em Michigan Mathematical Journal},
No. 6, 1959 , pp. 59-63.





\bibitem[GJ79]{GJ79} 
M.R. Garey and  D.S. Johnson,
{\em Computers and Intractability. A Guide to the Theory of NP-Completeness},
Freeman, San Francisco CA, 1979.





\bibitem[HK87]{HK87} 
A. Habel and H.J. Kreowski,
`` May we Introduce to you: Hypergraph Languages Generated by 
Hyperedge Replacement,'' 
{\em Proc. 13th International 
Workshop on Graph-Theoretic Concepts in Computer Science (WG'87) }, 
Springer Verlag, LNCS, Vol. 291, 1987, pp. 15-26.




\bibitem[HLW92]{HLW92} 
F. H\"ofting, T. Lengauer and E. Wanke,
``Processing of Hierarchically Defined Graphs and 
Graph Families,'' in
{\em Data Structures and Efficient Algorithms}
(Final Report on the DFG Special Joint Initiative),
Springer-Verlag, LNCS 594, 1992, pp. 44-69.


\bibitem[HW92]{HW92} 
F. H\"ofting and E. Wanke,
``Minimum Cost Paths in Periodic  Graphs,''
Tech Report 99, Universit \"at-Gesamthochschule-Paderborn, FRG,  
April, 1992. 



\bibitem[HU]{HU} 
J. E. Hopcroft, and J. D. Ullman,
{\em Introduction to Automata Theory, Languages and Computation}
Addison Wesley, Reading MA., 1979.




\bibitem[Hu73a]{Hu73a} H.B. Hunt III,
``On The Time Complexity of Languages,''
{\em Proc.  5th Annual ACM Symposium on Theory Of Computing, (STOC)}, 
1973, pp. 10-19.






\bibitem[IS87]{IS87}  K. Iwano and  K. Steiglitz,
``Testing for Cycles in  Infinite Graphs with Periodic Structure,''
{\em Proc.  19th Annual ACM Symposium on Theory of Computing, (STOC)}, 
1987, pp. 46-53.



\bibitem[IS88]{IS88}  K. Iwano and  K. Steiglitz,
``Planarity Testing of Doubly Connected Periodic Infinite Graphs,''
{\em Networks},
No. 18, 1988, pp. 205-222.




\bibitem[JL77]{JL77}
N.D.Jones and W.T.Laaser, 
``Complete Problems for Deterministic Polynomial Time,'' 
{\em Theoretical Computer Science},
No.  3, 1977, pp. 105-117.


\bibitem[KMW67]{KMW67} 
R.M. Karp, R.E. Miller and S. Winograd,
``The Organization of Computations for Uniform Recurrence Equations,''
{\em Journal of the ACM (J. ACM)}, Vol. 14, No. 3, 1967, pp. 563-590.




\bibitem[Ka72]{Ka72} 
R.M. Karp,
``Reducibility Among Combinatorial Problems,''
in R.E. Miller and J.W. Thatcher (eds)
{\em Complexity of Computer  Computations}, Plenum Press,
N.Y. 1972, pp. 85-103.





\bibitem[KO91]{KO91} 
M. Kodialam and  J.B. Orlin,
``Recognizing Strong Connectivity in Periodic graphs
and its relation to integer programming,''
{\em Proc. 2nd  ACM-SIAM Symposium on Discrete Algorithms (SODA)},
1991, pp. 131-135.





\bibitem[KS88]{KS88}  K. R. Kosaraju and G.F. Sullivan,
``Detecting Cycles in  Dynamic Graphs in Polynomial Time,''
{\em Proc.  27th IEEE Symposium on Foundations of Computer Science (FOCS)}, 
1988, pp. 398-406.




\bibitem[Le82]{Le82} 
T. Lengauer,
``The Complexity of  Compacting Hierarchically specified layouts of
Integrated circuits,''
{\em Proc. 23rd IEEE Symposium on Foundations of Computer Science (FOCS)}, 
1982, pp. 358-368.


\bibitem[LW87b]{LW87b} 
T. Lengauer and  C. Weiner,
``Efficient Solutions Hierarchical Systems of Linear Equations,''  
{\em  Computing}, Vol 39, 1987, pp. 111-132.




\bibitem[Le86]{Le86} 
T. Lengauer,
``Exploiting Hierarchy in VLSI Design,''
{\em Proc.  AWOC '86}, 
Springer-Verlag, LNCS 227, 1986, pp. 180-193.
	
\bibitem[LW87a]{LW87a} 
T. Lengauer and  E. Wanke,
``Efficient Solutions for Connectivity Problems for Hierarchically  
Defined Graphs ,''
{\em SIAM Journal on Computing}, 
Vol. 17, No. 6, 1988, pp. 1063-1080.




\bibitem[Le88]{Le88} 
T. Lengauer,
``Efficient Algorithms for Finding Minimum Spanning Forests of 
Hierarchically Defined graphs,''
{\em Journal of Algorithms}, Vol. 8, 1987, pp. 260-284.


\bibitem[Le89]{Le89} 
T. Lengauer,
``Hierarchical Planarity Testing,''
{\em Journal of the ACM (J. ACM) }, 
Vol. 36, No. 3, July 1989, pp. 474-509.


\bibitem[Le90]{Le90} 
T. Lengauer,
{\em Combinatorial Algorithms for Integrated Circuit Layout,}
John Wiley and Sons, 1990.


\bibitem[LW92]{LW92} 
T. Lengauer and  K.W. Wagner,
``The Correlation Between the Complexities of Non-Hierarchical and
Hierarchical Versions of Graph Problems,''
{\em Journal of Computer and System Sciences (JCSS)}, 
Vol. 44,  1992, pp. 63-93. 



\bibitem[LW93]{LW93} 
T. Lengauer and  E. Wanke,
``Efficient Decision Procedures for Graph Properties on Context-Free 
Graph Languages ,''
{\em Journal of the ACM (J. ACM) }, 
Vol. 40, No. 2, 1993, pp. 368-393.



\bibitem[Li82]{Li82} D. Lichtenstein,
``Planar Formulae and their Uses'', 
{\em SIAM Journal on Computing},
Vol 11, No. 2, May 1982 , pp. 329-343.




\bibitem[MH+93a]{MH+93a} M.V. Marathe H.B. Hunt III,  and S.S. Ravi,
``The Complexity of Approximating PSPACE-Complete Problems for 
Hierarchical Specifications,''
{\em Proc. 20th 
International Colloquium on Automata Languages and Programming (ICALP)}, 
July, 1993, pp. 76-87.


\bibitem[MR+93]{MR+93} 
M.V. Marathe, V. Radhakrishnan,  H.B. Hunt III,  and S.S. Ravi,
``Hierarchical Specified Unit Disk Graphs,''
{\em Proc. 19th International 
Workshop on Graph-Theoretic Concepts in Computer Science (WG '93)}, 
June, 1993, pp. 21-32. 




\bibitem[MH+94]{MH+94}  M.V. Marathe, H.B. Hunt III, 
R.E. Stearns and V. Radhakrishnan,
``Approximation schemes for
PSPACE-Complete problems for succinct graphs,'' 
{\em Proceedings of 26th Annual ACM
Symposium on the Theory of Computing (STOC),} May 1994,
pp 468-477.




\bibitem[Ma94]{Ma94} M.V. Marathe
{\em Complexity and Approximability of NP- and PSPACE-hard Optimization
Problems,}
Ph.D. thesis, Department of Computer Science, 
University at Albany,  Albany, NY August, 1994.


\bibitem[MH+95a]{MH+95a} M.V. Marathe, H.B. Hunt III and R.E. Stearns 
``Time as  Second Dimension: Complexity of Periodically Specified Problems,''
manuscript under preparation,
February, 1995.





\bibitem[MTM92]{MTM92} J.O. McClain, L.J. Thomas and J.B. Mazzola,
{\em Operations Management,}
Prentice Hall, Englewood Cliffs, 1992.

\bibitem[MC80]{MC80} C. Mead and L. Conway,
{\em Introduction to VLSI systems}
Addison Wesley, 1980.




\bibitem[MS81]{MS81} B.Monien and I.H.Sudborough, 
``Bounding the Bandwidth of NP-Complete Problems,''
{\em Proc. 13th ACM Annual Symposium on Theory of Computing (STOC)},
1981, pp. 279-292.






\bibitem[Or82a]{Or82a} J.B. Orlin,
``The Complexity of Dynamic/Periodic Languages and Optimization Problems,''
Sloan W.P. No. 1679-86 July 1985,
Working paper, Alfred P. Sloan School of Management,
MIT, Cambridge, MA 02139. A Preliminary version of the paper appears in
{\em Proc. 13th ACM Annual Symposium on Theory of Computing (STOC)}, 
1978, pp. 218-227.





\bibitem[Or82b]{Or82b} J.B. Orlin,
``Minimizing the Number of Vehicles to Meet a Fixed Periodic Schedule:
An Application to Periodic Posets,''
{\em Operations Research,}
No. 30, 1982, pp. 760-776.




\bibitem[Or83a]{Or83a} J.B. Orlin,
``Dynamic Matching and Quasidynamic Fractional Matchings I,''
{\em Networks}
Vol. 13,  1983, pp. 551-562.


\bibitem[Or83b]{Or83b} J.B. Orlin,
``Dynamic Matching and Quasidynamic Fractional Matchings II,''
{\em Networks}
Vol. 13,  1983, pp. 563-580.


\bibitem[Or84a]{Or84a} J.B. Orlin,
``Maximum Convex Cost  Dynamic Network Flows,''
{\em Mathematics for Operations Research,}
Vol. 9, No. 2, May 1984, pp. 190-206.


\bibitem[Or84b]{Or84b} J.B. Orlin,
``Some Problems on Dynamic/Periodic Graphs,''
{\em Progress in Combinatorial Optimization},
Academic Press, May 1984, pp. 273-293.



\iffalse******
\bibitem[PY91]{PY91} C. Papadimitriou and M. Yannakakis,
``Optimization, Approximation and Complexity Classes,''
{\em Journal of Computer and System Sciences (JCSS)}, 
No. 43, 1991, pp. 425-440.
************
\fi


\bibitem[Pa94]{Pa94} 
C. Papadimitriou,
{\em Computational Complexity}
Addison-Wesley, Reading, Massachusetts, 1994.




\bibitem[Ra93]{rven} V. Radhakrishnan,
{\em Sequential and Parallel Algorithms for Treewidth Bounded, Planar
and Hierarchically Specified Combinatorial Problems},
 Ph.D. thesis, Department of Computer Science, 
University at Albany,  Albany, NY August, 1993.


	
\bibitem[RH93]{RH93} D.J. Rosenkrantz and H.B. Hunt III,
``The Complexity of Processing Hierarchical Specifications'',
{\em SIAM Journal on Computing}, 
Vol. 22, No. 3, 1993, pp. 627-649.





\bibitem[Sa70]{Sa70}
W.J.Savitch, 
``Relationships Between Nondeterministic and Deterministic 
Tape Complexities,'' {\em Journal of Computer and System Sciences (JCSS),} 
Vol. 4, No. 2, pp. 177-192, 1970.


\bibitem[Sc78]{Sc78} 
T. Schaefer,
``The Complexity of Satisfiability Problems,''
{\em Proc. 10th ACM Symposium on Theory of Computing (STOC)}, 
1978, pp. 216-226.


\bibitem[Sc78a]{Sc78a} 
T. Schaefer,
``On the Complexity of Some Two-Person Perfect Information Games,''
{\em Journal of Computer and System Sciences (JCSS),} 
1978, pp. 185-225.



\bibitem[Sc76]{Sc76} C. Schnorr
``Satisfiability is quasilinear complete for NQL,''
{\em Journal of the ACM (J. ACM),} 
No. 25, 1976, pp. 136-145.


	
\bibitem[SH90]{SH90} R. E. Stearns and H. B. Hunt III
``Power Indices and Easier Hard Problems,''
{\em Mathematical Systems Theory}
Vol. 23, 1990, pp. 209-225.

\bibitem[Wa84]{Wa84}  K.W. Wagner,
``The  Complexity of Problems Concerning Graphs with Regularities'',
{\em Proc. 11th Symposium on Math. Foundations of Computer Science (MFCS)},
LNCS 176, Springer-Verlag, 1984, pp. 544-552.


\bibitem[Wa86]{Wa86} K.W. Wagner,
``The Complexity of Combinatorial Problems with Succinct 
Input Representation,''
{\em Acta Informatica }, Vol. 23, No. 3, 1986, pp. 325-356.



\bibitem[Wa93]{Wa93}  E.  Wanke,
``Paths and Cycles in Finite Periodic Graphs,''
{\em Proc. 20th Symposium on Math. Foundations of Computer Science (MFCS)},
LNCS 711, Springer-Verlag, 1993, pp. 751-760.


	
\bibitem[Wi90]{Wi90} M. Williams,
``Efficient Processing of  Hierarchical Graphs ,''
 {\em TR 90-06}, Dept of Computer Science, Iowa Sate University. 
(Parts of the report appeared in WADS'89, pp. 563-576
and SWAT'90, pp. 320-331 coauthored with 
Fernandez-Baca.)

\end{thebibliography}
